---
title: "605-Wk13-Discussion"
author: "Jose Mawyin"
date: "11/23/2019"
output: pdf_document
header-includes:
 \usepackage{float}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.pos = 'H')
library(mosaicCalc)
library(magrittr)
```
## 605-Wk13-Discussion

### 7.1 Area Between Curves Problem 7.10

\includegraphics[width=0.7\columnwidth]{/Users/josemawyin/Library/Mobile Documents/com~apple~CloudDocs/Data Science Masters /605/Calculus-7-10.png}


```{r}
F = antiD(sin(x)-cos(x) ~ x)
F
Answer7.10 = (F(x = 5*pi/4) - F(x = pi/4))
cat("The area under the curve is",Answer7.10 %>% round(2), "units.") 
```








\newpage

###  7.3 The Shell Method Problem 7.10


\includegraphics[width=0.7\columnwidth]{/Users/josemawyin/Downloads/ShellMethodCalculus-7-10.png}


\includegraphics[width=0.7\columnwidth]{/Users/josemawyin/Downloads/ShellMethodCalculus-7-10-Graph.jpeg}

```{r}
F = antiD(2*pi*(y^2)/5 ~ y)
F
Answer7.10 = (F(y = 10) - F(y = 5)) 
cat("The volume of the revolved solid is",Answer7.10 %>% round(2), "units.") 
```











