---
title: "605-Wk11-Homework"
author: "Jose Mawyin"
date: "11/9/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

### Linear regression model of stopping distance as a function of speed from the "Cars" dataset

First, lets load the dataset cars that contains 50 observations of two variables. The independent variable is vehicle's speed and the response variable is the vehicle's stopping distance.

```{r}
data("cars")
head(cars)
```

We can easily create a linear fit regression model.

```{r}
attach(cars)
Cars.lm <- lm(dist ~  speed)
```

Plotting a linear regression fit line we can see the speed vs. stopping distance response.

```{r,fig.height=3}
ggplot(cars, aes(x=speed, y=dist)) + geom_point() + geom_smooth(method='lm')
```

Looking into the statistics summary of the linear regression, we see that speed variable explains 65% of our data and that is has a P value (1.49e-12) well below the typical treshold of 0.05. 

```{r}
summary(Cars.lm)
```

The residuals plot below show how the residuals are evenly distributed around the origin (residual = 0). We can visualize this using a linear fit line that in our case passes directly over the x-axis as shown below.
```{r,fig.height=3}
ggplot(cars, aes(x=fitted(Cars.lm), y=resid(Cars.lm))) + geom_point() + geom_smooth(method='lm')
```

The residuals quantile-versus-quantile or Q-Q plot shows interesting results in that:

*...if the model fits the data well, we would expect the residuals to be normally (Gaussian) distributed around a mean of zero.*

We can see that in our Q-Q plot below that the residuals diverge around the tails indicating that the residuals are not normally distributed.


```{r }

qqnorm(resid(Cars.lm))
qqline(resid(Cars.lm))
```

We can see this more clearly below in a density plot of the residuals. We can easily see how the distribution of the residuals is positively skewed.


```{r,fig.height=3}
ggplot(cars, aes(resid(Cars.lm))) + geom_density(kernel = "gaussian") 
```





