---
title: "605-Wk14-HW"
author: "Jose Mawyin"
date: "12/3/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Ryacas)
library(tictoc)
library(mosaicCalc)
library(rSymPy)
library(NLP)
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

## Taylor Functions

| Function and Series   | First Few Terms   | Interval of Convergence   |
|--------------------------------------|--------------------------|------------------|
|   $\ln (1+x)=\sum_{n=1}^{\infty}(-1)^{(n-1)} \frac{x^{2 n-1}}{(2 n-1) !}$  |  $x-\frac{x^{2}}{2}+\frac{x^{3}}{3}-\frac{x^{4}}{4}+\frac{x^{5}}{5}-\dots$  |    $(-1,1]$   |
|  $e^{x}=\sum_{n=0}^{\infty} \frac{x^{n}}{n !}$  |  $1+x+\frac{x^{2}}{2 !}+\frac{x^{3}}{3 !}+\cdots$ |   $(-\infty, \infty)$   |
|    $\frac{1}{1-x}=\sum_{n=0}^{\infty} x^{n}$  |    $1+x+x^{2}+x^{3}+\cdots$ |     $(-1,1)$   |


For each function, only consider its valid ranges as indicated in the notes when you are computing the Taylor Series expansion. Please submit your assignment as a R-Markdown document.

We can use the library *rSymPy* to symbolically calculate the Taylor expansio of the series above. Below we compare how closely the Taylor expansion value compares to the numerical value evaluated at 0.8

$$\frac{1}{1-x}$$

```{r,tidy=TRUE, tidy.opts=list(width.cutoff=60)}
for (n in 1:10) {
sympy("var('p')")
sympy("var('x')") # or sympy("x = Symbol('x', real=True)")
equation <- paste("p=series(1/(1-x), x, 0,", n,")") #"p=series(1/(1-x), x, 0, 10)"
xt <- sympy(equation) # expand about 0 to 10th order
#xt <- sympy("p=series(1/(1-x), x, 0, 10)") # expand about 0 to 10th order
# Remove order information
xt0 <- sympy("p.removeO()")
# Test results
x <- 1/3
T1 <- eval(parse(text=xt0)) # Evaluate the result, xt0
T2 <- 1/(1-x)                # The correct value
Difference <- (T1-T2)
cat("\nAt", n," terms, The Taylor expansion is",xt0 ,". \n Evaulating at 0, the difference between the Taylor expansion and the built-in function is\n", Difference,"\n")
rm(x)
}
```

\pagebreak

$$e^{x}$$
```{r,tidy=TRUE, tidy.opts=list(width.cutoff=60)}
for (n in 1:10) {
sympy("var('p')")
sympy("var('x')") # or sympy("x = Symbol('x', real=True)")
equation <- paste("p=series(exp(x), x, 0,", n,")") #"p=series(1/(1-x), x, 0, 10)"
xt <- sympy(equation) # expand about 0 to 10th order
#xt <- sympy("p=series(1/(1-x), x, 0, 10)") # expand about 0 to 10th order
# Remove order information
xt0 <- sympy("p.removeO()")
# Test results
x <- 0.8
T1 <- eval(parse(text=xt0)) # Evaluate the result, xt0
T2 <- exp(x)                # The correct value
Difference <- (T1-T2)
cat("\nAt", n," terms, The Taylor expansion is",xt0 ,". \n Evaulating at 0.8, the difference between the Taylor expansion and the built-in function is\n", Difference,"\n")
rm(x)
}
```

\pagebreak

$$\ln (1+x)$$

```{r,tidy=TRUE, tidy.opts=list(width.cutoff=60)}
for (n in 1:10) {
sympy("var('p')")
sympy("var('x')") # or sympy("x = Symbol('x', real=True)")
equation <- paste("p=series(log(1+x), x, 0,", n,")") #"p=series(1/(1-x), x, 0, 10)"
xt <- sympy(equation) # expand about 0 to 10th order
#xt <- sympy("p=series(1/(1-x), x, 0, 10)") # expand about 0 to 10th order
# Remove order information
xt0 <- sympy("p.removeO()")
# Test results
x <- 0.8
T1 <- eval(parse(text=xt0)) # Evaluate the result, xt0
T2 <- log(1+x)                # The correct value
Difference <- (T1-T2)
cat("\nAt", n," terms, The Taylor expansion is",xt0 ,". \n Evaulating at 0.8, the difference between the Taylor expansion and the built-in function is\n", Difference,"\n")
rm(x)
}
```



