---
title: "605 Discussion Wk 6"
author: "Jose Mawyin"
date: "10/5/2019"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

## Chapter 4. Conditional Probability

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
perm = function(n, x) {
  factorial(n) / factorial(n-x)
}

comb = function(n, x) {
  factorial(n) / factorial(n-x) / factorial(x)
}
```

###15.  (a) What is the probability that your bridge partner has exactly two aces, given that she has at least one ace?
$$P\left( F|E \right) =\frac { P\left( F\cap E \right)  }{ P\left( E \right)  }$$
There are 52 different cards in a deck and each bridge hand has 13 cards. First we find the probability of having two aces given having at least one ace:
```{r}
Top.A <- (comb(48,11)*comb(4,2)/comb(52,13))
```

$$P\left( F\cap E \right) =\frac { \left( \begin{matrix} 48 \\ 11 \end{matrix} \right) \left( \begin{matrix} 4 \\ 2 \end{matrix} \right)  }{ \left( \begin{matrix} 52 \\ 13 \end{matrix} \right)  }$$


Then we find the probability of having at least once ace:
```{r}
Bottom.A <- (comb(52,13)-comb(48,13))/comb(52,13)
```
$$P\left( E \right) =\frac { \left( \begin{matrix} 52 \\ 13 \end{matrix} \right) -\left( \begin{matrix} 48 \\ 13 \end{matrix} \right)  }{ \left( \begin{matrix} 52 \\ 13 \end{matrix} \right)  } $$
Therefore:
```{r}
Top.A/Bottom.A
```

$$P\left( F|E \right) =\frac { P\left( F\cap E \right)  }{ P\left( E \right)  } =\frac { \left( \begin{matrix} 48 \\ 11 \end{matrix} \right) \left( \begin{matrix} 4 \\ 2 \end{matrix} \right)  }{ \left( \begin{matrix} 52 \\ 13 \end{matrix} \right) -\left( \begin{matrix} 48 \\ 13 \end{matrix} \right)  } =0.307$$


###     (b) What is the probability that your bridge partner has exactly two aces, given that she has the ace of spades?
$$P\left( F|E \right) =\frac { P\left( F\cap E \right)  }{ P\left( E \right)  }$$

Just as before, we start knowing that there are 52 different cards in a deck and each bridge hand has 13 cards. We need to find first the probability:
```{r}
Top.B <- (comb(48, 11)*comb(3, 1))/comb(52,13)
```

$$P\left( F\cap E \right) =\frac { \left( \begin{matrix} 48 \\ 11 \end{matrix} \right) \left( \begin{matrix} 3 \\ 1 \end{matrix} \right)  }{ \left( \begin{matrix} 52 \\ 13 \end{matrix} \right)  } $$

To find the the probability of having the ace of spades:
```{r}
Bottom.B <- comb(51,12)/comb(52,13)
```

$$P\left( E \right) =\frac { \left( \begin{matrix} 51 \\ 12 \end{matrix} \right)  }{ \left( \begin{matrix} 52 \\ 13 \end{matrix} \right)  }$$


Therefore:
```{r}
Top.B/Bottom.B
```
$$P\left( F|E \right) =\frac { P\left( F\cap E \right)  }{ P\left( E \right)  } =\frac { \left( \begin{matrix} 48 \\ 11 \end{matrix} \right) \left( \begin{matrix} 3 \\ 1 \end{matrix} \right)  }{ \left( \begin{matrix} 51 \\ 12 \end{matrix} \right)  } =0.427$$



