---
title: "605 - HW 5"
author: "Jose Mawyin"
date: "9/25/2019"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(dplyr)
library(Ryacas)
```

## Choose independently two numbers B and C at random from the interval [0, 1] with uniform density. Prove that B and C are proper probability distributions.
Note that the point (B,C) is then chosen at random in the unit square.

###Find the probability that:
```{r}
B <- runif(10000, 0, 1)
C <- runif(10000, 0, 1)
#B;C
#hist(A); hist(B)
```


```{r}

```


###(a) B + C < 1/2.
```{r}
#(B + C) < 1/2
SUM.B.C <- (B + C)
hist(SUM.B.C, main = "Histogram of Distribution of B + C", col = "yellow", breaks=20)
```

```{r}
length(which(SUM.B.C < 1/2))/length(SUM.B.C)
```
**The probability that B + C < 1/2 approaches 0.13**
```{r}

```


###(b) BC < 1/2.
```{r}
#B*C < 1/2
B.Times.C <- B*C
hist(B.Times.C, col = "yellow", breaks=20)
length(which(B.Times.C < 1/2))/length(B.Times.C)
```
**The probability that BC < 1/2 approaches 0.85**

###(c) |B − C| < 1/2.
```{r}
#abs(B - C) < 1/2
Abs.B.minus.C <- abs(B - C)
hist(Abs.B.minus.C, col = "yellow", breaks=20)
length(which(Abs.B.minus.C < 1/2))/length(Abs.B.minus.C)
```
**The probability that |B − C| < 1/2 approaches 0.75**

###(d) max{B,C} < 1/2.
```{r}
#max{B,C}
max.B.C <- pmax(B,C)
hist(max.B.C, col = "yellow", breaks=20)
length(which(max.B.C < 1/2))/length(max.B.C)
```
**The probability that max{B,C} < 1/2 approaches 0.25**

###(e) min{B,C} < 1/2.
```{r}
min.B.C <- pmin(B,C)
hist(min.B.C, col = "yellow", breaks=20)
length(which(min.B.C < 1/2))/length(min.B.C)
```
**The probability that min{B,C} < 1/2 approaches 0.75**
 
 